Actuator line modeling of vertical-axis turbines

### 2016-05-12 paper.tex
Remove RM2 wake figs

Figure~\ref{fig:RM2-ALM-meancontquiv} shows the mean velocity field at 1 m
downstream or $x/D=0.93$ computed by the ALM for the RM2. The mean near-wake
structure looks similar to the RVAT ALM RANS case, but for the RM2, the lack of
detail from blade tip vortex shedding matches more closely with
experiments~\cite{Bachant2016-RM2-paper}. However, the vertical flow towards the
$x$--$y$ center plane was captured, which is an important qualitative feature of
both CFT near-wakes.

    \includegraphics[width=0.75\textwidth]{RM2-ALM_meancontquiv}

    \caption{Mean velocity field at $x/D=0.93$ for the RM2 predicted by the
        ALM.}

    \label{fig:RM2-ALM-meancontquiv}
\end{figure}

Figure~\ref{fig:RM2-ALM-kcont} shows the ALM's turbulence kinetic energy
predictions in the near-wake of the RM2. Like for the UNH-RVAT, $k$ appears to
be concentrated on the $+y$ side of the rotor. However, overall levels of
turbulence are lower than for the UNH-RVAT, which is consistent with the
experimental results. However, turbulence generated by the RM2's blade tip
vortex shedding was not captured by the ALM.

    \includegraphics[width=0.7\textwidth]{RM2-ALM_kcont}

    \caption{Turbulence kinetic energy contours at $x/D=0.93$ behind the RM2
        predicted by the ALM.}

    \label{fig:RM2-ALM-kcont}
\end{figure}

Wake profiles at turbine mid-height in the $x$--$y$ center plane are shown in
Figure~\ref{fig:RM2-ALM-profiles}. Like the UNH-RVAT case, the mean velocity
deficit appears to be recovering too quickly, which may similarly be due to
coarseness of the grid or overprediction of the eddy viscosity. The peak in
turbulence kinetic energy on the $-y$ side of the turbine was also
underpredicted, though was much lower in magnitude compared to the UNH-RVAT,
even in the experimental measurements.

    \includegraphics[width=0.7\textwidth]{RM2-ALM_wake-profiles}

    \caption{Mean streamwise velocity (left) and turbulence kinetic energy
        (right) profiles at $x/D=0.93$ and $z/H=0$ computed for the RM2 using the
        ALM inside a $k$--$\epsilon$ RANS model, compared with experimental data
        from~\cite{Bachant2016-RM2-data}.}

    \label{fig:RM2-ALM-profiles}
\end{figure}

Finally, a similar mean streamwise momentum transport analysis was undertaken by
computing weighted sums of each term across the entire domain in the $y$--$z$
directions. Similar results as for the UNH-RVAT were obtained, i.e.,
cross-stream advection was predicted to be positive where it should have been
negative, vertical advection was predicted reasonably well, and turbulent
transport due to the eddy viscosity was also relatively large. The ratio of wake
transport compared with the UNH-RVAT RANS case (approximately 60\% to 70\% lower
for the RM2) matches well with that computed from the experiments, which shows
the ALM may successfully predict larger optimal array spacing for the RM2 versus
UNH-RVAT.

    \includegraphics[width=0.7\textwidth]{RM2-ALM_recovery-bar-chart}

    \caption{Weighted average momentum recovery terms at $x/D=0.93$ for the RM2
        actuator line model with a $k$--$\epsilon$ RANS closure and the experiments
        described in \cite{Bachant2016-RM2-paper}.}

    \label{fig:RM2-ALM-recovery}
\end{figure}

>>>>

The mean near-wake structure for the RM2 looked qualitatively similar to the
RVAT ALM RANS case, but for the RM2, the effects of blade tip vortex shedding
were weaker, which is consistent with experiments~\cite{Bachant2016-RM2-paper}.
Nonetheless, the mean vertical flow towards the $x$--$y$ center plane was
captured, which is an important qualitative feature of both CFT near-wakes.
Like for the UNH-RVAT, predicted turbulence kinetic energy values were
concentrated on the $+y$ side of the rotor. However, overall levels of
turbulence are lower than for the UNH-RVAT, which is also consistent with the
experimental results.

As for the UNH-RVAT RANS case, a mean streamwise momentum transport analysis was
undertaken for the RM2 near-wake by computing weighted sums of each term across
the entire domain in the $y$--$z$ directions. Similar results as for the
UNH-RVAT were obtained, i.e., cross-stream advection was predicted to be
positive where it should have been negative, vertical advection was predicted
reasonably well, and turbulent transport due to the eddy viscosity was also
relatively large. The ratio of wake transport compared with the UNH-RVAT RANS
case (approximately 60\% to 70\% lower for the RM2) matches well with that
computed from the experiments, which shows the ALM may successfully predict
larger optimal array spacing for the RM2 versus UNH-RVAT.
